---
title: "Retail Analysis"
author: "--"
date: "`r Sys.Date()`"
output: html_document
---

```{r, echo=FALSE}
# Install and load necessary libraries
library(tidyverse)
library(caret)
library(lubridate)
library(dplyr)
library(corrplot)
library(knitr)
library(ggplot2)
```

```{r, echo=FALSE}
# Data Collection
sales_data <- read.csv(file.choose())
features_data <- read.csv(file.choose())
stores_data <- read.csv(file.choose())
```

```{r, echo=FALSE}
# Data Cleaning
clean_sales_data <- sales_data %>%
  filter(Weekly_Sales >= 0)# Handling negative values in Sales data frame

clean_features_data <- features_data %>%
  drop_na() %>%
  distinct()
clean_stores_data <- stores_data %>%
  drop_na() %>%
  distinct()
```

```{r, echo=FALSE}
# Convert the Date column to a Date format in all data frames
clean_sales_data$Date <- as.Date(clean_sales_data$Date, format="%d/%m/%Y")
clean_features_data$Date <- as.Date(clean_features_data$Date, format="%d/%m/%Y")
```

```{r, echo=FALSE}
# Merge the datasets using left joins
merged_data <- sales_data %>%
  left_join(features_data, by = c("Store", "Date")) %>%
  left_join(stores_data, by = "Store")

# Print the head of the merged data
head(merged_data)
```

# Figure 7. Holiday Reference for Corresponding Week Numbers

```{r, echo=FALSE}
# Create the Week column
merged_data$Week <- lubridate::week(merged_data$Date)
# Merge the holiday_table with the lookup table using IsHoliday as the k

# Convert the Date column to a Date format
merged_data$Date <- as.Date(merged_data$Date, format="%d/%m/%Y")

# Create the Week column
merged_data$Week <- lubridate::week(merged_data$Date)

# Create the Holiday_Ref column based on the specified holiday names
merged_data$Holiday_Ref <- ifelse(merged_data$Week == 6, "SupperBowl Week",
                                  ifelse(merged_data$Week == 36, "LaborDay Week",
                                         ifelse(merged_data$Week == 47, "ThanksGiving Week",
                                                ifelse(merged_data$Week == 52, "Christmas Week", "Non Holiday Week"))))

# Print the updated merged_data
head(merged_data)
```

# Figure 7. Holiday Reference with their respecive weeks
```{r, echo=FALSE}
# Filter merged_data to include only rows with holidays
holiday_week_ref <- merged_data %>%
  filter(Holiday_Ref != "Non Holiday Week") %>%
  select(Week, Holiday_Ref) %>%
  distinct()
kable(holiday_week_ref)
```
Fig. 7. Holiday Reference for Corresponding Week Numbers

# Figure 8. Weekely sales vs Stores

```{r, echo=FALSE}
# Weekly_Sales vs. Store
ggplot(merged_data, aes(x = Store, y = Weekly_Sales, color = Holiday_Ref)) +
  geom_point() +
  labs(title = "Weekly Sales vs. Store",
       x = "Store",
       y = "Weekly Sales") +
  scale_x_continuous(breaks = seq(0, 45, by = 1), limits = c(0, 45))
```

The analysis of weekly sales across various stores reveals a prominent surge in sales during Thanksgiving Week (Week 47), suggesting a notable correlation between this holiday period and heightened purchasing activity. Additionally, a general upward trend in holiday sales is observed across most stores, indicating increased customer spending during these periods. While Thanksgiving Week stands out as a high-sales period, store-specific variations in performance during other holidays highlight opportunities for tailored marketing and strategic promotions. The findings suggest the importance of understanding both overarching trends and individual store dynamics for effective sales strategies.

# Figure 9. Weekely sales vs Temperature
```{r, echo=FALSE}
# Weekly_Sales vs. Temperature
ggplot(merged_data, aes(x = Temperature, y = Weekly_Sales, color = Holiday_Ref)) +
  geom_point() +
  labs(title = "Weekly Sales vs. Temperature",
       x = "Temperature",
       y = "Weekly Sales") +
  scale_x_continuous(breaks = seq(-3, 105, by = 2), limits = c(-3, 105))
```

The examination of the weekly sales in relation to temperature patterns revealed noteworthy aspects into consumer behavior. The majority of sales happens within a temperature range of 19 to 79 degrees. This indicate a preference for moderate climatic conditions. Further, holidays such as Christmas, SuperBowl, and ThanksGiving showed distinct temperature-sales correlations which saw Christmas and ThanksGiving having increased sales in the range of 17 to 65 degrees, which aligns with a preference for cooler weather during these festivities. The Labor Day, conversely, shows elevated sales in the warmer range of 55 to 90 degrees. Moreover, ThanksGiving proves to be a pivotal driver of high sales as compared to sales in other holidays, emphasizing the significance of strategic planning and targeted marketing efforts during this holiday. This analysis gives the importance of considering holiday dynamics and temperature variations in crafting effective sales strategies.
